---
title: "Simulation Set up for ReMPCA"
output: pdf_document
---

- 1. Sparsity on *v* (1-SE)
- 2. Smoothness on *v*
- 3. Smoothness and Sparsity on *v* 
- 4.1. Sparsity on *u* (1-SE)
- 4.2. Smoothness on *u*
- 5. Two-way conditional sparsity
- 6. Two-way conditional smoothness
- 7. Conditional two-way sparsity two-way smoothness
- 8. Multivariate hybrid data case



## Sparsity on v
Starting with the first case, we can do it for a simulated functional data:
```{r load-functions, include=FALSE}
source("Sparsity.R")
source("Smoothness.R")
source("fdClass.R")
source("rdClass.R")
source("hdClass.R")
source("Power_Algorithm.R")
source("get_pen.R")
```


Simulated data:
```{r}
x <- seq(0,2*pi, len = 50)
y1 <- sin(x); y1[1:26] <- 0 
y2 <- cos(2*x); y2[32:50] <- 0
z<- outer(y1,y2)
  
par(mfrow = c(1,3))
plot(x,y1, type = 'l', ylim = range(y1,y2))
points(x,y2, type = 'l', col='red')

image(z)
persp(z, theta = 40)

set.seed(123)
X1 <- X2 <- data.frame(matrix(nrow = 100, ncol = length(y1)))
for (i in 1:150) {
  X1[i,] <- y1 + rnorm(length(y1), sd = 0.5)
  X2[i,] <- y2 + rnorm(length(y2), sd = 0.5)
}

par(mfrow = c(1,2))
matplot(t(X1), type ='l', main ='Variable 1')
matplot(t(X2), type ='l', main ='Variable 2')
```

Sparsity: 
```{r, include=FALSE}
fd_object <- fdClass(data = as.matrix(X1),
                     argval = NULL, 
                     Smoothing_parameter = NULL,  
                     Sparsity_parameter = round(seq(0,49, length.out = 30)))  

rd_object <- rdClass(data = as.matrix(X2),
                     Sparsity_parameter = round(seq(0,49, length.out = 30)))


hd_list <- list(fd_object, rd_object)  # List of fd and rd objects
object_list <- hdClass(hdlist = hd_list,
                       argval = NULL,  # Grid points for rows
                       Smoothing_parameter = NULL,  # Custom smoothing parameter for rows
                       Sparsity_parameter = round(seq(0,100, length.out = 50)))  # Custom sparsity parameter for rows

# Combine matrices side by side
hd <- object_list
n <- nrow(hd); K_fold <- 5; sparse_tuning_type <- "soft"
n_var <- attr(object_list, "n_var") # Number of variables (# of matrices in object_list)
ncol <- attr(object_list, "ncol") # Number of columns of each matrix

####### Smoothing Parameter ##########
# Generate all combinations alphas (one row per combination)
smooth_tuning_col <- expand.grid(attr(object_list, "Smoothing_parameter_col")) # A matrix
smooth_tuning_row <- attr(object_list, "Smoothing_parameter") # A vector

####### level of sparsity (for both functional and non-functional data) #######
sparsity_row_list <- attr(object_list, "Sparsity_parameter")
sparsity_col_list <- attr(object_list, "Sparsity_parameter_col")

X_temp <- hd
# No smoothness list:
S_alpha_v <- list()
for (i in 1:n_var) {
  S_alpha_v[[i]] <- diag(ncol[,i])
  }

    
CV_score_sparse_u <- CV_score_sparse_v <- GCV_score_smooth_u <- GCV_score_smooth_v <- Inf
CV_scores_result_u <- CV_scores_result_v <- c()

gamma_v <- rep(0, n_var)
for (i in 1:n_var) {
  gamma_Xi <- sparsity_col_list[[i]]
  CV_scores_result_v <- c()
  for (sparse_tuning_single in gamma_Xi) {
    gamma_v[i] <- sparse_tuning_single
    sparse_score = cv_sparse_col(data = X_temp,
                                 n_var = n_var,
                                 ncol = ncol,
                                 S_alpha_v = S_alpha_v,
                                 S_alpha_u = diag(nrow(X_temp)),
                                 K_fold = K_fold,
                                 sparse_tuning_result_u = 0, #gamma_u,
                                 sparse_tuning_result_v = gamma_v,
                                 sparse_tuning_type = sparse_tuning_type)

    CV_scores_result_v <- c(CV_scores_result_v, sparse_score)
    if (sparse_score <= CV_score_sparse_v) {
      CV_score_sparse_v = sparse_score
      sparse_tuning_selection_v = sparse_tuning_single
    }
    gamma_v[i] <- gamma_Xi[which.min(CV_scores_result_v)]
  }
}
```

By implementing cross validation, we'll get the following result:
```{r}
par(mfrow = c(2,2))
matplot(t(X1), type = 'l', main = 'X1')
matplot(t(X2), type = 'l', main = 'X2')
gamma_v

result <-power_algo(data = X_temp,
                    n_var = n_var,
                    ncol = ncol,
                    sparse_tuning_result_u = 0, # A fixed number
                    sparse_tuning_result_v = gamma_v, # A vector (length = p)
                    S_alpha_v = S_alpha_v, # A list (lenght = p)
                    S_alpha_u = diag(nrow(X_temp)), # A matrix
                    sparse_tuning_type= sparse_tuning_type)

v <- result[[1]]
matplot(-v[1:50], type = 'l', main = 'v1')
matplot(-v[51:100], type = 'l', main = 'v2')
```
By implementing 1-SE approach, we can see the results below: 
```{r, include=FALSE}
cv_sparse_col_SE <- function(data,
                             n_var,
                             ncol,
                             S_alpha_v,
                             S_alpha_u,
                             K_fold,
                             sparse_tuning_result_u,
                             sparse_tuning_result_v,
                             sparse_tuning_type) {

  set.seed(123)
  shuffled_rows <- sample(nrow(data)) # Grouping rows of data matrix
  group_size <- ifelse(round(length(shuffled_rows) / K_fold, digits = 0) == 0,
                       1,
                       round(length(shuffled_rows) / K_fold, digits = 0))

  data_tilde <- data
  fold_errors <- numeric(K_fold)  # store errors

  for (k in 1:K_fold) {
    rows_to_remove <- shuffled_rows[((k - 1) * group_size + 1):min(k * group_size, length(shuffled_rows))]
    data_train <- data.frame(data_tilde[-rows_to_remove,])
    data_test <- data.frame(matrix(data_tilde[rows_to_remove,], nrow = length(rows_to_remove)))

    # Power Algorithm
    power_train <- power_algo(data = data_train,
                              n_var = n_var,
                              ncol = ncol,
                              sparse_tuning_result_u = sparse_tuning_result_u,
                              sparse_tuning_result_v = sparse_tuning_result_v,
                              S_alpha_v = S_alpha_v,
                              S_alpha_u = S_alpha_u[-rows_to_remove, -rows_to_remove],
                              sparse_tuning_type = sparse_tuning_type)

    v_train <- power_train[[1]]
    u_test <- as.matrix(data_test) %*% as.matrix(v_train)

    # Fold-wise error
    fold_errors[k] <- sum((data_test - u_test %*% t(v_train))^2) / nrow(data_test)
  }
  CV_mean <- mean(fold_errors)
  CV_se <- sd(fold_errors) / sqrt(K_fold)
  
  return(list(CV_error = CV_mean, SE = CV_se, fold_errors = fold_errors))
}

# Tuning across sparsity levels using 1-SE rule
CV_score_sparse_u <- CV_score_sparse_v <- GCV_score_smooth_u <- GCV_score_smooth_v <- Inf
CV_scores_result_u <- CV_scores_result_v <- list()
CV_results <- list()

gamma_v <- rep(0, n_var)
for (i in 1:n_var) {
  gamma_Xi <- sparsity_col_list[[i]]
  cv_means <- cv_ses <- fold_errors_vec <- numeric(length(gamma_Xi))

  for (j in seq_along(gamma_Xi)) {
    gamma_v[i] <- gamma_Xi[j]

    cv_result <- cv_sparse_col_SE(data = X_temp,
                                  n_var = n_var,
                                  ncol = ncol,
                                  S_alpha_v = S_alpha_v,
                                  S_alpha_u = diag(nrow(X_temp)),
                                  K_fold = K_fold,
                                  sparse_tuning_result_u = 0,
                                  sparse_tuning_result_v = gamma_v,
                                  sparse_tuning_type = sparse_tuning_type)

    cv_means[j] <- cv_result$CV_error
    cv_ses[j] <- cv_result$SE
    #fold_errors_vec <- cv_result$fold_errors
  }
  
  CV_results[[i]] <- data.frame(gamma_Xi, cv_means, cv_ses)
  
  # 1-SE rule
  j_min <- which.min(cv_means)
  threshold <- cv_means[j_min] + cv_ses[j_min]
  j_1se <- max(which(cv_means <= threshold))  # Most sparse within 1-SE

  # Store results
  CV_scores_result_v[[i]] <- list(errors = cv_means, SEs = cv_ses)
  sparse_tuning_selection_v <- gamma_Xi[j_1se]
  gamma_v[i] <- sparse_tuning_selection_v
}

gamma_v
```

```{r}
par(mfrow = c(2,2))
matplot(t(X1), type = 'l', main = 'X1')
matplot(t(X2), type = 'l', main = 'X2')
gamma_v

result <-power_algo(data = X_temp,
                    n_var = n_var,
                    ncol = ncol,
                    sparse_tuning_result_u = 0, # A fixed number
                    sparse_tuning_result_v = gamma_v, # A vector (length = p)
                    S_alpha_v = S_alpha_v, # A list (lenght = p)
                    S_alpha_u = diag(nrow(X_temp)), # A matrix
                    sparse_tuning_type= sparse_tuning_type)

v <- result[[1]]
matplot(-v[1:50], type = 'l', main = 'v1')
matplot(-v[51:100], type = 'l', main = 'v2')
```



## Smoothness on v
Now, let's implement smoothness on v and use GCV to tune the parameters. This time, we will assume that all alpha values for the second variable (regular data) are zero. The formula of GCV is as follow:
$$ \text{GCV}(\alpha) = \frac{1}{m} \sum_{k = 1}^K \frac{\lVert (I-\pmb{S}^{[{\alpha}]}_k)(\pmb{X}_k^\top u) \rVert^2}{ (1 - \frac{1}{m_k}\text{tr}(\pmb{S}^{[{\alpha}]}_k))^2}$$
```{r, include=FALSE}
smoothness_type <- "Second_order"
smooth_tuning_col <- expand.grid(attr(object_list, "Smoothing_parameter_col")) # A matrix

####### Grid Points #######
GridPoints_u <- attr(object_list, "GridPoints_u")
GridPoints_v <- attr(object_list, "GridPoints_v")

####### S_alpha for all alphas #######
S_alpha_list_u <- S_alpha_list_v <- Omegas_v <- list()
index <- 0
cat("Preprocessing ...\n")
n_iter1 <- nrow(smooth_tuning_col)     # The number of alphas
pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                     max = n_iter1,# Maximum value of the progress bar
                     style = 3,    # Progress bar style (also available style = 1 and style = 2)
                     width = 50,   # Progress bar width. Defaults to getOption("width")
                     char = "=")   # Character used to create the bar

# S_alpha for v
for (alpha_index in 1:nrow(smooth_tuning_col)) {
  index <- index + 1
  S_u <- S_v <- Omegas_v <- Omegas_u <- list()
  for (i in 1:n_var) {
    alpha <- as.numeric(smooth_tuning_col[alpha_index,i])
    if(is.null(GridPoints_v[[i]])){
      S_v[[i]] <- diag(ncol[,i])
    }else{
      get.pen.result <- get.pen(td = GridPoints_v[[i]],
                                alpha = alpha,
                                type = smoothness_type)
      S_v[[i]] <- get.pen.result$S.alpha
      Omegas_v[[i]] <- get.pen.result$ Omega
    }
  }
  S_alpha_list_v[[index]] <- S_v #as.matrix(bdiag(S_v))
  setTxtProgressBar(pb, index)
}


opt_alpha_v <- function(X,
                        n_var,
                        ncol,
                        S_alphas_v,  # List of length = n_iter, each element is list of length = n_var
                        S_alphas_u,  # Matrix (n × n)
                        alphas,      # Matrix of alpha_v values (n_iter × n_var)
                        alpha_u,
                        sparse_tuning_result_u,
                        sparse_tuning_result_v,
                        Omega_u,
                        sparse_tuning_type) {

  n_iter <- nrow(alphas)
  n <- nrow(X)
  m <- ncol(X)
  GCV <- numeric(n_iter)

  if (all(alphas == 0)) {
    return(list(GCV_v = Inf,
                opt.alpha_v = rep(0, n_var),
                opt_s.alpha_v = replicate(n_var, diag(m / n_var), simplify = FALSE),
                GCVdf_v = data.frame(alphas, rep(Inf, n_iter))))
  } else {
    for (i in 1:n_iter) {
      S_alpha_v <- S_alphas_v[[i]]  # list of S_k
      GCV_alpha <- 0
      trace_total <- 0

      power_result <- power_algo(data = X,
                                 n_var = n_var,
                                 ncol = ncol,
                                 sparse_tuning_result_u = sparse_tuning_result_u,
                                 sparse_tuning_result_v = sparse_tuning_result_v,
                                 S_alpha_v = S_alpha_v,
                                 S_alpha_u = S_alphas_u,
                                 sparse_tuning_type = sparse_tuning_type)

      u <- power_result[[2]]  # fixed left singular vector
      df <- X

      for (k in 1:n_var) {
        m_k <-as.integer(ncol[k])
        X_k <- df[, 1:m_k]
        df <- df[, -(1:m_k), drop = FALSE]
        S_k <- S_alpha_v[[k]]

        # Residual norm after smoothing
        residual <- (diag(m_k) - S_k) %*% (t(X_k) %*% u)
        GCV_alpha <- GCV_alpha + (1 / m_k) * (norm_vec(residual)^2)
        trace_total <- trace_total + sum(diag(S_k))
      }

      # Final GCV with global denominator
      GCV[i] <- GCV_alpha / (1 - (trace_total / m))^2
    }

    opt.alpha <- alphas[which.min(GCV), ]
    opt_s.alpha <- S_alphas_v[[which.min(GCV)]]

    return(list(GCV_v = GCV,
                opt.alpha_v = opt.alpha,
                opt_s.alpha_v = opt_s.alpha,
                GCVdf_v = data.frame(alphas, GCV)))
  }
}


GCV_result_v = opt_alpha_v(X = X_temp,
                           n_var = n_var,
                           ncol = ncol,
                           S_alphas_v = S_alpha_list_v, # A list (length = n_var)
                           S_alphas_u = diag(nrow(X_temp)),
                           alphas = smooth_tuning_col, # A matrix
                           alpha_u = 0,
                           sparse_tuning_result_u = 0,
                           sparse_tuning_result_v = c(0,0),
                           Omega_u = diag(nrow(X_temp)),
                           sparse_tuning_type = sparse_tuning_type)
```

```{r}
GCV_result_v$opt.alpha_v
par(mfrow = c(2,2))
matplot(t(X1), type = 'l', main = 'X1')
matplot(t(X2), type = 'l', main = 'X2')

result <-power_algo(data = X_temp,
                    n_var = n_var,
                    ncol = ncol,
                    sparse_tuning_result_u = 0, # A fixed number
                    sparse_tuning_result_v = c(0,0), # A vector (length = p)
                    S_alpha_v = GCV_result_v$opt_s.alpha_v, # A list (lenght = p)
                    S_alpha_u = diag(nrow(X_temp)), # A matrix
                    sparse_tuning_type= sparse_tuning_type)

v_ss <- result[[1]]
matplot(-v_ss[1:50], type = 'l', main = 'v1')
matplot(-v_ss[51:100], type = 'l', main = 'v2')

par(mfrow = c(1,1))
plot(GCV_result_v$GCV_v, type = 'l')
```

Another case is when both variables are considered functional. In this scenario, we will tune the smoothness parameters for both variables.

```{r, include=FALSE}
fd_object <- fdClass(data = as.matrix(X1),
                     argval = NULL, 
                     Smoothing_parameter = NULL,  
                     Sparsity_parameter = round(seq(0,49, length.out = 30)))  

fd_object2 <- fdClass(data = as.matrix(X2),
                     argval = NULL, 
                     Smoothing_parameter = NULL,
                     Sparsity_parameter = round(seq(0,49, length.out = 30)))


hd_list <- list(fd_object, fd_object2)  # List of fd and rd objects
object_list <- hdClass(hdlist = hd_list,
                       argval = NULL,  # Grid points for rows
                       Smoothing_parameter = NULL,  # Custom smoothing parameter for rows
                       Sparsity_parameter = round(seq(0,100, length.out = 50)))  # Custom sparsity parameter for rows

smoothness_type <- "Second_order"
smooth_tuning_col <- expand.grid(attr(object_list, "Smoothing_parameter_col")) # A matrix

####### Grid Points #######
GridPoints_u <- attr(object_list, "GridPoints_u")
GridPoints_v <- attr(object_list, "GridPoints_v")

####### S_alpha for all alphas #######
S_alpha_list_u <- S_alpha_list_v <- list()
index <- 0
cat("Preprocessing ...\n")
n_iter1 <- nrow(smooth_tuning_col)     # The number of alphas
pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                     max = n_iter1,# Maximum value of the progress bar
                     style = 3,    # Progress bar style (also available style = 1 and style = 2)
                     width = 50,   # Progress bar width. Defaults to getOption("width")
                     char = "=")   # Character used to create the bar

# S_alpha for v
for (alpha_index in 1:nrow(smooth_tuning_col)) {
  index <- index + 1
  S_u <- S_v <- Omegas_v <- Omegas_u <- list()
  for (i in 1:n_var) {
    alpha <- as.numeric(smooth_tuning_col[alpha_index,i])
    if(is.null(GridPoints_v[[i]])){
      S_v[[i]] <- diag(ncol[,i])
    }else{
      get.pen.result <- get.pen(td = GridPoints_v[[i]],
                                alpha = alpha,
                                type = smoothness_type)
      S_v[[i]] <- get.pen.result$S.alpha
      Omegas_v[[i]] <- get.pen.result$ Omega
    }
  }
  S_alpha_list_v[[index]] <- S_v #as.matrix(bdiag(S_v))
  setTxtProgressBar(pb, index)
}

GCV_result_v = opt_alpha_v(X = X_temp,
                           n_var = n_var,
                           ncol = ncol,
                           S_alphas_v = S_alpha_list_v, # A list (length = n_var)
                           S_alphas_u = diag(nrow(X_temp)),
                           alphas = smooth_tuning_col, # A matrix
                           alpha_u = 0,
                           sparse_tuning_result_u = 0,
                           sparse_tuning_result_v = c(0,0),
                           Omega_u = diag(nrow(X_temp)),
                           sparse_tuning_type = sparse_tuning_type)
```

```{r}
GCV_result_v$opt.alpha_v
par(mfrow = c(2,2))
matplot(t(X1), type = 'l', main = 'X1')
matplot(t(X2), type = 'l', main = 'X2')

result <-power_algo(data = X_temp,
                    n_var = n_var,
                    ncol = ncol,
                    sparse_tuning_result_u = 0, # A fixed number
                    sparse_tuning_result_v = c(0,0), # A vector (length = p)
                    S_alpha_v = GCV_result_v$opt_s.alpha_v, # A list (lenght = p)
                    S_alpha_u = diag(nrow(X_temp)), # A matrix
                    sparse_tuning_type= sparse_tuning_type)

v_ss <- result[[1]]
matplot(-v_ss[1:50], type = 'l', main = 'v1')
matplot(-v_ss[51:100], type = 'l', main = 'v2')

par(mfrow = c(1,1))
plot(GCV_result_v$GCV_v, type ='l')
```
## Smoothness and Sparsity on *v*
Next, we can have both sparsity and smoothness on v, using the tuned parameters in the previous step:

```{r, fig.height=10, fig.width=6}
par(mfrow = c(3,2))
matplot(t(X1), type = 'l', main = 'X1')
matplot(t(X2), type = 'l', main = 'X2')
matplot(y1, type = 'l', main = 'Variable 1')
matplot(y2, type = 'l', main = 'Variable 2')

result <-power_algo(data = X_temp,
                    n_var = n_var,
                    ncol = ncol,
                    sparse_tuning_result_u = 0, # A fixed number
                    sparse_tuning_result_v = gamma_v, # A vector (length = p)
                    S_alpha_v = GCV_result_v$opt_s.alpha_v, # A list (lenght = p)
                    S_alpha_u = diag(nrow(X_temp)), # A matrix
                    sparse_tuning_type= sparse_tuning_type)

v_ss <- result[[1]]
matplot(-v_ss[1:50], type = 'l', main = 'v1', lwd = 2)
points(y1/5, type = "l", col = 'red', lty = 3, lwd = 0.7)
matplot(-v_ss[51:100], type = 'l', main = 'v2', lwd = 2)
points(y2/5.1, type = "l", col = 'red', lty = 3, lwd = 0.7)
```

If we use the following formula for conditional GCV when $\alpha_u = 0$, we should get the same result:


$$\text{GCV}_v ({\alpha}_v; \alpha_u = 0) =
\frac{1}{m} \sum_{k = 1}^K
\frac{ \left\| \frac{\pmb{X}_k^\top u}{\|u\|^2} - v_k \right\|^2 }
{\left( 1 - \frac{1}{m_k} \cdot \frac{ \operatorname{tr}(\pmb{S}_v(\alpha_k)) }{ 1 + \alpha_u \cdot \mathcal{R}_u(u) } \right)^2}$$

```{r}
power_algo2 = function(data,
                      n_var,
                      ncol,
                      sparse_tuning_result_u, # A fixed number
                      sparse_tuning_result_v, # A vector (length = p)
                      S_alpha_v, # A list (lenght = p)
                      S_alpha_u, # A matrix
                      sparse_tuning_type,
                      conditional = FALSE, # New parameter for conditional update
                      Omega_u = NULL,      # Needed only if conditional = TRUE
                      Omega_v = NULL,      # Needed only if conditional = TRUE
                      alpha_v = 0,  # A vector
                      alpha_u = 0   # A number  
                      ){

  rownames(data) <- NULL; colnames(data) <- NULL
  data <- as.matrix(data)

  # Splitting data
  start_col <- 1; splitted_data <- list()
  for (i in 1:n_var) {
    num_cols <- ncol[[i]]  # Get number of columns for the ith matrix
    splitted_data[[i]] <- data[, start_col:(start_col + num_cols - 1)]  # Extract matrix
    start_col <- start_col + num_cols
  }

  v_old <- svd(as.matrix(data))$v[, 1]
  errors <- Inf; thresh <- 1e-10

  # Power Algorithm
  while (errors > thresh) {
    u_new <- S_alpha_u %*% sparse_pen_fun(y = data%*%as.matrix(v_old),
                                          tuning_parameter = sparse_tuning_result_u,
                                          type = sparse_tuning_type) # u = h_{gamma} Xv
    if(all(u_new == 0)){ # To avoid 0 in the denominator
      u_new <- S_alpha_u %*% data%*%v_old
      }
    v_new <- c()
    for(i in 1:n_var){
      sparse_param <- as.integer(sparse_tuning_result_v[i])
      v <- S_alpha_v[[i]] %*% sparse_pen_fun(y = t(splitted_data[[i]]) %*% as.matrix(u_new),
                                             tuning_parameter = sparse_param,
                                             type = sparse_tuning_type) # v = S_{alpha} h_{gamma} t(X)u

      if(all(v == 0)){ # To avoid 0 in the denominator
        v <- S_alpha_v[[i]] %*% t(splitted_data[[i]]) %*% as.matrix(u_new)}
      v_new <- c(v_new, v)
      }

    v_new = v_new / norm_vec(v_new) # v/||v||

    # Convergence condition
    errors = sum((v_new - v_old)^2)
    v_old <- v_new
  }
  u_new = u_new/norm_vec(u_new) # u/||u||
  
  if (conditional && !is.null(Omega_u) && !is.null(Omega_v)) {
    u_new <- u_new/(t(v_new) %*% (diag(nrow(Omega_v)) + alpha_v * Omega_v) %*% v_new)
    v_new <- v_new/(t(u_new) %*% (diag(nrow(Omega_u)) + alpha_u * Omega_u) %*% u_new)
  }
  return(list(v_new,u_new))
}



opt_alpha_v <- function(X,
                        n_var,
                        ncol,
                        S_alphas_v,   # List of length = n_iter, each element is list of length = n_var
                        S_alphas_u,   # Matrix (n × n)
                        alphas,       # Matrix of alpha_v values (n_iter × n_var)
                        alpha_u,      # scalar
                        sparse_tuning_result_u,
                        sparse_tuning_result_v,
                        Omega_u,
                        sparse_tuning_type) {

  n_iter <- nrow(alphas)
  n <- nrow(X)
  m <- ncol(X)
  GCV <- numeric(n_iter)

  if (all(alphas == 0)) {
    return(list(GCV_v = Inf,
                opt.alpha_v = rep(0, n_var),
                opt_s.alpha_v = replicate(n_var, diag(m / n_var), simplify = FALSE),
                GCVdf_v = data.frame(alphas, rep(Inf, n_iter))))
  } else {
    for (i in 1:n_iter) {
      S_alpha_v <- S_alphas_v[[i]]  # list of S_k for current alpha set
      GCV_alpha <- 0

      # Get u and v from power algorithm
      power_result <- power_algo2(data = X,
                                  n_var = n_var,
                                  ncol = ncol,
                                  sparse_tuning_result_u = sparse_tuning_result_u,
                                  sparse_tuning_result_v = sparse_tuning_result_v,
                                  S_alpha_v = S_alpha_v,
                                  S_alpha_u = S_alphas_u,
                                  sparse_tuning_type = sparse_tuning_type,
                                  conditional = TRUE,
                                  Omega_u = Omega_u,
                                  Omega_v = Omega_v,
                                  alpha_v = alpha_v,
                                  alpha_u = alpha_u)
      
      

      v <- power_result[[1]]  # full v vector
      u <- power_result[[2]]  # fixed u

      norm_u2 <- as.numeric(norm_vec(u)^2)
      Xu_proj <- as.vector(t(X) %*% u / norm_u2)  # (X' u / ||u||^2)
      R_u <- as.numeric(t(u) %*% Omega_u %*% u / norm_u2)

      idx <- 0
      for (k in 1:n_var) {
        m_k <- as.integer(ncol[k])
        v_k <- v[(idx + 1):(idx + m_k)]
        xuk <- Xu_proj[(idx + 1):(idx + m_k)]
        S_k <- S_alpha_v[[k]]

        denom <- (1 - (1 / m_k) * (sum(diag(S_k)) / (1 + alpha_u * R_u)))^2
        numer <- (1 / m_k) * norm_vec(xuk - v_k)^2

        GCV_alpha <- GCV_alpha + numer / denom
        idx <- idx + m_k
      }

      GCV[i] <- GCV_alpha / n_var  # Optional: averaging; or keep as total across variables
    }

    opt.alpha <- alphas[which.min(GCV), ]
    opt_s.alpha <- S_alphas_v[[which.min(GCV)]]

    return(list(GCV_v = GCV,
                opt.alpha_v = opt.alpha,
                opt_s.alpha_v = opt_s.alpha,
                GCVdf_v = data.frame(alphas, GCV)))
  }
}



GCV_result_v = opt_alpha_v(X = X_temp,
                           n_var = n_var,
                           ncol = ncol,
                           S_alphas_v = S_alpha_list_v, # A list (length = n_var)
                           S_alphas_u = diag(nrow(X_temp)),
                           alphas = setNames(data.frame(unique(smooth_tuning_col[,1]) ,
                                                        rep(0, length(unique(smooth_tuning_col[,1])))), c("alpha1", "alpha2")),
                           alpha_u = 1,
                           sparse_tuning_result_u = 0,
                           sparse_tuning_result_v = c(0,0),
                           Omega_u = diag(nrow(X_temp)),
                           sparse_tuning_type = sparse_tuning_type)


GCV_result_v$opt.alpha_v
GCV_result_v$GCVdf_v
```


