---
title: "Simulation Set up for ReMPCA"
output: pdf_document
---

1.	Sparsity on v (Including Confidence Interval)
2.	Smoothness on v
3. Smoothness and Sparsity on v (Conditional-order)
4_1.	Sparsity on u (Including Confidence Interval)
4_2.	Smoothness on u
5.	Two-way conditional sparsity 
6.	Two-way conditional smoothness 
7.	Conditional two-way sparsity two-way smoothness 
8.	Multivariate hybrid data case 


## Sparsity on v
Starting with the first case, we can do it for a simulated functional data:
```{r load-functions, include=FALSE}
source("Sparsity.R")
source("Smoothness.R")
source("fdClass.R")
source("rdClass.R")
source("hdClass.R")
source("Power_Algorithm.R")
```

```{r, include=FALSE}
K_fold = 5
sparse_tuning_type = "soft"
sigma1 = 1; sigma2 = 0.5

d = 100                   # Length of each sin signal
t = seq(0, 1, length.out = d)  # Generate 'd' points from 0 to 1
nn = 300                  # Total number of observations (e.g., rows)
t_total = d + d           # Here t_total = 200, i.e., two segments each of length d

v11 = sin(1 * pi * c(t, rep(0, d)))
v12 = sin(2 * pi * c(rep(0, d), t))
U111 <- matrix(rnorm(nn/3, mean = 2, sd = 0.1), ncol=1)
U121 <- matrix(rnorm(nn/3, mean = 1, sd = 0.01), ncol=1)
U131 <- matrix(rnorm(nn/3, mean = 0, sd = 0.1), ncol=1)
U112 <- matrix(rnorm(nn/3, mean = 0, sd = 0.1), ncol=1)
U122 <- matrix(rnorm(nn/3, mean = 1, sd = 0.1), ncol=1)
U132 <- matrix(rnorm(nn/3, mean = 2, sd = 0.1), ncol=1)

# Combine them into a (nn) x 2 matrix
U1_obj = rbind(
  cbind(U111),
  cbind(U121),
  cbind(U131)
)

U2_obj = rbind(
  cbind(U112),
  cbind(U122),
  cbind(U132)
)


### 5. Construct the data matrix X1:

X1 = U1_obj %*% 10 %*% rbind(v11) +
  matrix(
    rnorm(t_total * nn, mean = 0, sd = sigma1),
    nrow = nn
  )

X2= U2_obj %*% 3 %*% rbind(v12) +
  matrix(
    rnorm(t_total * nn, mean = 0, sd = sigma2),
    nrow = nn
  )
fd_object <- fdClass(data = X1,
                     argval = NULL,  # Grid points for columns
                     Smoothing_parameter = NULL,  # Custom smoothing parameter
                     Sparsity_parameter = round(seq(0,199, length.out = 30)))  # Custom sparsity parameter

rd_object <- rdClass(data = X2,
                     Sparsity_parameter = round(seq(0,199, length.out = 30)))


hd_list <- list(fd_object, rd_object)  # List of fd and rd objects
object_list <- hdClass(hdlist = hd_list,
                       argval = NULL,  # Grid points for rows
                       Smoothing_parameter = NULL,  # Custom smoothing parameter for rows
                       Sparsity_parameter = round(seq(0,300, length.out = 30)))  # Custom sparsity parameter for rows

# Combine matrices side by side
hd <- object_list
n <- nrow(hd)
n_var <- attr(object_list, "n_var") # Number of variables (# of matrices in object_list)
ncol <- attr(object_list, "ncol") # Number of columns of each matrix

####### Smoothing Parameter ##########
# Generate all combinations alphas (one row per combination)
smooth_tuning_col <- expand.grid(attr(object_list, "Smoothing_parameter_col")) # A matrix
smooth_tuning_row <- attr(object_list, "Smoothing_parameter") # A vector

####### level of sparsity (for both functional and non-functional data) #######
sparsity_row_list <- attr(object_list, "Sparsity_parameter")
sparsity_col_list <- attr(object_list, "Sparsity_parameter_col")

X_temp <- hd
# No smoothness list:
S_alpha_v <- list()
for (i in 1:n_var) {
  S_alpha_v[[i]] <- diag(ncol[,i])
  }

    
CV_score_sparse_u <- CV_score_sparse_v <- GCV_score_smooth_u <- GCV_score_smooth_v <- Inf
CV_scores_result_u <- CV_scores_result_v <- c()

gamma_v <- rep(0, n_var)
for (i in 1:n_var) {
  gamma_Xi <- sparsity_col_list[[i]]
  CV_scores_result_v <- c()
  for (sparse_tuning_single in gamma_Xi) {
    gamma_v[i] <- sparse_tuning_single
    sparse_score = cv_sparse_col(data = X_temp,
                                 n_var = n_var,
                                 ncol = ncol,
                                 S_alpha_v = S_alpha_v,
                                 S_alpha_u = diag(nrow(X_temp)),
                                 K_fold = K_fold,
                                 sparse_tuning_result_u = 0, #gamma_u,
                                 sparse_tuning_result_v = gamma_v,
                                 sparse_tuning_type = sparse_tuning_type)

    CV_scores_result_v <- c(CV_scores_result_v, sparse_score)
    if (sparse_score <= CV_score_sparse_v) {
      CV_score_sparse_v = sparse_score
      sparse_tuning_selection_v = sparse_tuning_single
    }
    gamma_v[i] <- gamma_Xi[which.min(CV_scores_result_v)]
  }
}
```

By considering a low sigma (1 for var1 and 0.5 for var2), we'll get the following results:
```{r}
par(mfrow = c(2,2))
matplot(t(X1), type = 'l', main = 'X1')
matplot(t(X2), type = 'l', main = 'X2')
gamma_v

result <-power_algo(data = X_temp,
                    n_var = n_var,
                    ncol = ncol,
                    sparse_tuning_result_u = 0, # A fixed number
                    sparse_tuning_result_v = gamma_v, # A vector (length = p)
                    S_alpha_v = S_alpha_v, # A list (lenght = p)
                    S_alpha_u = diag(nrow(X_temp)), # A matrix
                    sparse_tuning_type= sparse_tuning_type)

v <- result[[1]]
matplot(v[1:200], type = 'l', main = 'v1')
matplot(v[201:400], type = 'l', main = 'v2')
```
Using 1-SE rule in sparsity on v:
```{r}
cv_sparse_col <- function(data,
                          n_var,
                          ncol,
                          S_alpha_v,
                          S_alpha_u,
                          K_fold,
                          sparse_tuning_result_u,
                          sparse_tuning_result_v,
                          sparse_tuning_type) {

  set.seed(123)
  shuffled_rows <- sample(nrow(data)) # Grouping rows of data matrix
  group_size <- ifelse(round(length(shuffled_rows) / K_fold, digits = 0) == 0,
                       1,
                       round(length(shuffled_rows) / K_fold, digits = 0))

  data_tilde <- data
  fold_errors <- numeric(K_fold)  # store errors

  for (k in 1:K_fold) {
    rows_to_remove <- shuffled_rows[((k - 1) * group_size + 1):min(k * group_size, length(shuffled_rows))]
    data_train <- data.frame(data_tilde[-rows_to_remove,])
    data_test <- data.frame(matrix(data_tilde[rows_to_remove,], nrow = length(rows_to_remove)))

    # Power Algorithm
    power_train <- power_algo(data = data_train,
                              n_var = n_var,
                              ncol = ncol,
                              sparse_tuning_result_u = sparse_tuning_result_u,
                              sparse_tuning_result_v = sparse_tuning_result_v,
                              S_alpha_v = S_alpha_v,
                              S_alpha_u = S_alpha_u[-rows_to_remove, -rows_to_remove],
                              sparse_tuning_type = sparse_tuning_type)

    v_train <- power_train[[1]]
    u_test <- as.matrix(data_test) %*% as.matrix(v_train)

    # Fold-wise error
    fold_errors[k] <- sum((data_test - u_test %*% t(v_train))^2) / nrow(data_test)
  }
  CV_mean <- mean(fold_errors)
  CV_se <- sd(fold_errors) / sqrt(K_fold)
  
  return(list(CV_error = CV_mean, SE = CV_se, fold_errors = fold_errors))
}

# Tuning across sparsity levels using 1-SE rule
CV_score_sparse_u <- CV_score_sparse_v <- GCV_score_smooth_u <- GCV_score_smooth_v <- Inf
CV_scores_result_u <- CV_scores_result_v <- list()
CV_results <- data.frame()

gamma_v <- rep(0, n_var)
for (i in 1:n_var) {
  gamma_Xi <- sparsity_col_list[[i]]
  cv_means <- cv_ses <- fold_errors_vec <- numeric(length(gamma_Xi))

  for (j in seq_along(gamma_Xi)) {
    gamma_v[i] <- gamma_Xi[j]

    cv_result <- cv_sparse_col(data = X_temp,
                               n_var = n_var,
                               ncol = ncol,
                               S_alpha_v = S_alpha_v,
                               S_alpha_u = diag(nrow(X_temp)),
                               K_fold = K_fold,
                               sparse_tuning_result_u = 0,
                               sparse_tuning_result_v = gamma_v,
                               sparse_tuning_type = sparse_tuning_type)

    cv_means[j] <- cv_result$CV_error
    cv_ses[j] <- cv_result$SE
    fold_errors_vec <- cv_result$fold_errors
  }
  
  # 1-SE rule
  j_min <- which.min(cv_means)
  threshold <- cv_means[j_min] + cv_ses[j_min]
  print(threshold)
  j_1se <- max(which(cv_means <= threshold))  # Most sparse within 1-SE

  # Store results
  CV_scores_result_v[[i]] <- list(errors = cv_means, SEs = cv_ses)
  sparse_tuning_selection_v <- gamma_Xi[j_1se]
  gamma_v[i] <- sparse_tuning_selection_v
}

gamma_v
```




Now, if we implement the smoothness on v using GCV, we'll get:













## Two-way data
```{r}
U1 <- function(s) sin(2 * pi * s)
U2 <- function(s) sin(2 * pi * (s - 0.25))
V1 <- function(t) -3 + 8 * exp(-4 * (t - 0.25)^2)
V2 <- function(t) -3 + 8 * exp(-4 * (t - 0.75)^2)

s <- seq(from = 0, to = 1, length.out = 201)
t <- seq(from = 0, to = 1, length.out = 201)

par(mfrow = c(2, 2))
ts.plot(U1(s), main = "U1"); ts.plot(U2(s), main = "U2")
ts.plot(V1(t), main = "V1"); ts.plot(V2(t), main = "V2")
x_star <- outer(U1(s),V1(t)) + outer(U2(s),V2(t)) #U1(s)*V1(t)+U2(s)*V2(t)
par(mfrow = c(1, 1))
persp(x_star, main = "x*")
```





Small Simulation
```{r}
x <- seq(0,2*pi, len = 50)
y1 <- sin(x); y1[1:26] <- 0 
y2 <- cos(2*x); y2[32:50] <- 0
z<- outer(y1,y2)
  
par(mfrow = c(1,3))
plot(x,y1, type = 'l', ylim = range(y1,y2))
points(x,y2, type = 'l', col='red')

image(z)
persp(z, theta = 40)

X1 <- X2 <- data.frame(matrix(nrow = 100, ncol = length(y1)))
for (i in 1:100) {
  X1[i,] <- y1 + rnorm(length(y1), sd = 0.5)
  X2[i,] <- y2 + rnorm(length(y2), sd = 0.5)
}

par(mfrow = c(1,2))
matplot(t(X1), type ='l', main ='Variable 1')
matplot(t(X2), type ='l', main ='Variable 2')
```

Sparsity: 
```{r}
fd_object <- fdClass(data = as.matrix(X1),
                     argval = NULL, 
                     Smoothing_parameter = NULL,  
                     Sparsity_parameter = round(seq(0,49, length.out = 30)))  

rd_object <- rdClass(data = as.matrix(X2),
                     Sparsity_parameter = round(seq(0,49, length.out = 30)))


hd_list <- list(fd_object, rd_object)  # List of fd and rd objects
object_list <- hdClass(hdlist = hd_list,
                       argval = NULL,  # Grid points for rows
                       Smoothing_parameter = NULL,  # Custom smoothing parameter for rows
                       Sparsity_parameter = round(seq(0,100, length.out = 50)))  # Custom sparsity parameter for rows

# Combine matrices side by side
hd <- object_list
n <- nrow(hd)
n_var <- attr(object_list, "n_var") # Number of variables (# of matrices in object_list)
ncol <- attr(object_list, "ncol") # Number of columns of each matrix

####### Smoothing Parameter ##########
# Generate all combinations alphas (one row per combination)
smooth_tuning_col <- expand.grid(attr(object_list, "Smoothing_parameter_col")) # A matrix
smooth_tuning_row <- attr(object_list, "Smoothing_parameter") # A vector

####### level of sparsity (for both functional and non-functional data) #######
sparsity_row_list <- attr(object_list, "Sparsity_parameter")
sparsity_col_list <- attr(object_list, "Sparsity_parameter_col")

X_temp <- hd
# No smoothness list:
S_alpha_v <- list()
for (i in 1:n_var) {
  S_alpha_v[[i]] <- diag(ncol[,i])
  }

    
CV_score_sparse_u <- CV_score_sparse_v <- GCV_score_smooth_u <- GCV_score_smooth_v <- Inf
CV_scores_result_u <- CV_scores_result_v <- c()

gamma_v <- rep(0, n_var)
for (i in 1:n_var) {
  gamma_Xi <- sparsity_col_list[[i]]
  CV_scores_result_v <- c()
  for (sparse_tuning_single in gamma_Xi) {
    gamma_v[i] <- sparse_tuning_single
    sparse_score = cv_sparse_col(data = X_temp,
                                 n_var = n_var,
                                 ncol = ncol,
                                 S_alpha_v = S_alpha_v,
                                 S_alpha_u = diag(nrow(X_temp)),
                                 K_fold = K_fold,
                                 sparse_tuning_result_u = 0, #gamma_u,
                                 sparse_tuning_result_v = gamma_v,
                                 sparse_tuning_type = sparse_tuning_type)

    CV_scores_result_v <- c(CV_scores_result_v, sparse_score)
    if (sparse_score <= CV_score_sparse_v) {
      CV_score_sparse_v = sparse_score
      sparse_tuning_selection_v = sparse_tuning_single
    }
    gamma_v[i] <- gamma_Xi[which.min(CV_scores_result_v)]
  }
}
```









